Choose a License

### 2024-12-13 choose_license.qmd
Add SIL OFL again and remove CC0
SIL OFL is the de facto standard license
(e.g., the only accepted license by Google
Fonts and recommended by Fedora), while CC0
is not even mentioned.

- __fonts:__ Copyleft licenses applied to fonts can be a special case: If a font is put under the license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), any documents containing texts using that font will probably be derivative works and have to be put under the same license if shared. As a non-copyleft license, [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) is a good choice for fonts.
- __templates and TeX packages:__ If a template or, for this purpose, a $\TeX$ package, is licensed under a copyleft license such as the (A)GPLv3, every work that is a derivative of the template or that uses the $\TeX$ package has to put under the same license, if shared [@Koppor2016]. And if a document contained source code covered by the (A)GPLv3, the same license would also apply to the document [@MadHatter2022].
- __database content:__ If the work produced from a database (the "output") is a derivative of the content in the database, the output is subject to the restrictions laid out in the license. For example, if geospatial data were to be licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), all maps produced from the data would likely need to fulfill this license's obligation for unrestricted access (its Anti-DRM provision) if shared [see @Poole2017]. Similarly, following an example from @Matt2009, if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to be licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) (or a compatible license) if shared. If the intention is to only have [derivative databases](https://osmfoundation.org/wiki/Licence/Community_Guidelines/Produced_Work_-_Guideline) under the same license, one might want to choose the [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/) for the database, as it was specifically designed not to apply to works produced from the data in the database. Otherwise [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) is an excellent choice for data.
- __works in the public domain__: If a work is already in the public domain, it should be marked using the [PDM 1.0](https://creativecommons.org/publicdomain/mark/1.0/), rather than applying a waiver such as the [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) (or another license).
:::

>>>>

- __fonts:__ Copyleft licenses applied to fonts can be a special case: If a font is put under the license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), any documents containing texts using that font will probably be derivative works and have to be put under the same license if shared. The [SIL Open Font License\ 1.1](https://openfontlicense.org/) is the _de facto_ standard license for fonts, requiring that derived fonts, if published, have to be put under the same license. It doesn't require attribution for usage (e.g., when embedded in a document), but forbids that the font is sold by itself. Selling the font is allowed, however, if it is bundled with other software.
- __templates and TeX packages:__ If a template or, for this purpose, a $\TeX$ package, is licensed under a copyleft license such as the (A)GPLv3, every work that is a derivative of the template or that uses the $\TeX$ package has to put under the same license, if shared [@Koppor2016]. And if a document contained source code covered by the (A)GPLv3, the same license would also apply to the document [@MadHatter2022].
- __database content:__ If the work produced from a database (the "output") is a derivative of the content in the database, the output is subject to the restrictions laid out in the license. For example, if geospatial data were to be licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), all maps produced from the data would likely need to fulfill this license's obligation for unrestricted access (its Anti-DRM provision) if shared [see @Poole2017]. Similarly, following an example from @Matt2009, if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to be licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) (or a compatible license) if shared. If the intention is to only have [derivative databases](https://osmfoundation.org/wiki/Licence/Community_Guidelines/Produced_Work_-_Guideline) under the same license, one might want to choose the [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/) for the database, as it was specifically designed not to apply to works produced from the data in the database. Otherwise [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) is an excellent choice for data.
- __works in the public domain__: If a work is already in the public domain, it should be marked using the [PDM 1.0](https://creativecommons.org/publicdomain/mark/1.0/), rather than applying a waiver such as the [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) (or another license).
:::
